Generating urban morphologies at large scales

### 2019-03-15 paper.tex
.

Simulation results and real measures are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/LGK0US}. Source code is available on the git repository of the project at \url{https://github.com/openmole/spatialdata}.
% TODO split repositories

@@

@@

%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \rotatebox{90}{Class 1}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_2.43416877400257_48.5494220403648_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_3.20479601880545_50.1572174642354_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_6.03454192624923_43.1438923676911_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_7.76412991205779_47.6437664789465_1}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 2}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_4.73115258198923_45.9763617017002_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_4.92601740592234E-5_50.7945845258906_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_8.76440233864255_50.3459471560717_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_10.6351277289413_51.1689134935896_2}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 3}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_2.78910884212325_50.5481512216399_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_20.0294237103859_51.5168560281418_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_5.55350495469112_51.4926898192205_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_23.8025598321731_54.950616933903_3}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 4}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_5.28978808584814_52.0605158929046_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_6.11373040780706_50.9165373258647_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_8.35334588815557_49.530653800312_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_12.4396942996003_50.983284784414_4}.png}
  \end{subfigure}\\
  \caption{Typology applied to OpenStreetMap samples. The first class correspond mostly to a high density of linear buildings, recalling modern urbanism projects; the second disparate large buildings that can corresponds to industrials buildings in the outskirt of cities; the third disparate small buildings closer to periurban settlements; the fourth are denser and more complicated patterns evoking historical urban centers.}
  % provide coordinates ?
  \label{fig:osm_types}
\end{figure}
%%%%%%%%%%%%%%

@@

A simulation experiment provides an insight into the patterns produced by the different generators in the morphological space. We sample the parameter space using a Latin Hypercube Sampling, with 10000 points for each generator respectively, and with 100 stochastic repetitions for each parameter point.

@@

To quantify the level of calibration of each generator regarding each centroid, and in average regarding stochastic repetitions, we provide in Table~

@@

Our approach provides a first step towards systematic modeling of generative processes of urban form at large scales, and can thus be extended in several ways.
First of all, we focused on the built environment but neglected transportation infrastructures, whereas spatial network morphogenesis models have been proposed for example by \cite{courtat2011mathematics} or \cite{raimbault2018multi} in a multi-modeling approach.
Taking into account multiple dimensions of the urban system is an important extension and hybrid models such as co-evolution models \citep{raimbault2014hybrid} should be investigated.

@@

%%%%%%%%%%%%%%%%%%%%%
%\section{Conclusion}

>>>>

Simulation results and real measures are available on the dataverse repository at \url{https://doi.org/10.7910/DVN/LGK0US}. Source code is available on the git repository of the project at \url{https://github.com/openmole/spatialdata}.
% TODO split repositories
The model and indicators were coded in scala langage for performance purposes. This furthermore allows a seamless integration into the OpenMOLE workflow engine for model exploration \citep{reuillon2013openmole}, which provides methods for numerical experiments (in our case sampling methods) and transparent access to high performance computation environments (model simulation were run on the European Grid Infrastructure for an equivalent of one year and one month CPU time).

@@

%%%%%%%%%%%%%%
\begin{figure*}
\vspace{-1cm}
  \centering
  \rotatebox{90}{Class 1}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/1/osm_2.43416877400257_48.5494220403648_1}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/1/osm_3.20479601880545_50.1572174642354_1}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/1/osm_6.03454192624923_43.1438923676911_1}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/1/osm_7.76412991205779_47.6437664789465_1}.png}}
  \end{subfigure}\\
  \rotatebox{90}{Class 2}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/2/osm_4.73115258198923_45.9763617017002_2}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/2/osm_4.92601740592234E-5_50.7945845258906_2}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/2/osm_8.76440233864255_50.3459471560717_2}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/2/osm_10.6351277289413_51.1689134935896_2}.png}}
  \end{subfigure}\\
  \rotatebox{90}{Class 3}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/3/osm_2.78910884212325_50.5481512216399_3}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/3/osm_20.0294237103859_51.5168560281418_3}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/3/osm_5.55350495469112_51.4926898192205_3}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/3/osm_23.8025598321731_54.950616933903_3}.png}}
  \end{subfigure}\\
  \rotatebox{90}{Class 4}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/4/osm_5.28978808584814_52.0605158929046_4}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/4/osm_6.11373040780706_50.9165373258647_4}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/4/osm_8.35334588815557_49.530653800312_4}.png}}
  \end{subfigure}
  \begin{subfigure}[b]{0.13\textwidth}
    \framebox{\includegraphics[width=\textwidth]{{osm/4/osm_12.4396942996003_50.983284784414_4}.png}}
  \end{subfigure}\\
  \caption{Typology applied to OpenStreetMap samples. The first class correspond mostly to a high density of linear buildings, recalling modern urbanism projects; the second disparate large buildings that can corresponds to industrials buildings in the outskirt of cities; the third disparate small buildings closer to periurban settlements; the fourth are denser and more complicated patterns evoking historical urban centers.}
  % provide coordinates ?
  \label{fig:osm_types}
\end{figure*}
%%%%%%%%%%%%%%

@@

@@

A simulation experiment provides an insight into the patterns produced by the different generators in the morphological space. We sample the parameter space using a Latin Hypercube Sampling, with 10000 points for each generator respectively, and with 100 stochastic repetitions for each parameter point. This sampling is achieved with scripting the models into the OpenMOLE platform~\citep{reuillon2013openmole}.

@@

%%%%%%%%%%%%%%%%%
\begin{table*}[h]
\center{
\begin{tabular}{|c|c|c|c|c|}\hline
           & Random & Blocks & Exp. Mixture & Percolation\\ \hline\hline
Centroid 1 & $0.424 \pm 0.011$ & $0.106 \pm 0.063$ & $0.303\pm 0.101$ & $0.325\pm 0.019$ \\
Centroid 2 &$0.809\pm 0.022$&$0.164\pm 0.099$&$0.184\pm 0.141$&$0.947\pm 0.019$\\
Centroid 3 &$0.428\pm 0.019$&$0.095\pm 0.054$&$0.109\pm 0.064$&$0.541\pm 0.019$\\
Centroid 4 &$0.515 \pm 0.005$&$0.311\pm 0.077$&$0.589\pm 0.149$ &$0.083\pm 0.025$\\\hline
\end{tabular}
}
\vskip 0.25cm
\caption{Aggregated distance in the morphological space, for each generator and each calibration objective (real clusters centroids). Euclidian distance in the projected space are aggregated in average on stochastic repetitions, and the minimal average value is reported with its standard deviation.\label{tab:averagedists}}
\end{table*}
%%%%%%%%%%%%%%%%%

To quantify the level of calibration of each generator regarding each centroid, and in average regarding stochastic repetitions, we provide in Table~\ref{tab:averagedists} the aggregated minimal values of distances, for each generator and each calibration objective, with their standard deviations. This mainly confirms the previous results, with however interesting variations: (i) for the second centroid, the exponential mixture is in average no longer the best, and furthermore has a higher variability; (ii) centroid three and four are the easiest to reach, despite the latest being in the boundary; (iii) the percolation generator performs well on this point and has a very low variability.

This experiment has therefore shown the possibility to calibrate the generative models on morphological measures against real configurations, and furthermore unveils their complementarity to approach the diverse existing forms.

@@

Our approach provides a first step towards systematic modeling of generative processes of urban form at large scales. Some direct limitations could be tackled in a short term. Testing slightly different processes and heuristics in generator may be a way to cover the part of the real point cloud which is missed by our generators. As it seems correspond to complicated urban centers, it may be however complicated without more elaborated models. Also, we did not use minimization algorithms to calibrate the generators, and the and a further step would consist in checking the robustness of our result using such optimization heuristics (genetic algorithms for example), but also diversity algorithms such as pattern space exploration proposed by \cite{10.1371/journal.pone.0138212}, to ensure the effective feasible space of each generator.

This work can also be extended in several ways. First of all, we focused on the built environment but neglected transportation infrastructures, whereas spatial network morphogenesis models have been proposed for example by \cite{courtat2011mathematics} or \cite{raimbault2018multi} in a multi-modeling approach.
Taking into account multiple dimensions of the urban system is an important extension and hybrid models such as co-evolution models \citep{raimbault2014hybrid} should be investigated.

@@

%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

We have proposed here a new insight into the generative simulation of urban morphologies at large scales, namely the scale of the district considering the layout of buildings. After computing morphological measures on a large sample of real urban areas, we showed the complementarity of different generators capturing various aspects of urban morphogenesis processes. Despite not implying generative agents (developers, inhabitants, companies) and thus staying close to procedural modeling, this work however paves the way towards a more systematic understanding of generative processes of urban form at this scale.
